RILS-ROLS: Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-10-26 doc/paper.tex
Exp. section: working

Note that when the algorithm terminates with $R^2=1$ and $RMSE=0$, the match of found model and the exact model is proved -- the results are reported in the second block. However, in case when $R^2 \neq 1$, that is $R^2 \geq 0.999$, we manually check these cases comparing the match of obtained model and the exact model, indicating the result. \fxnote{Aco, projeri ovu recenicu, da nisam promasio...}

@@

\begin{itemize}
	\item Concerning the exact results in case noise is not presented, the best performing algorithm is our \textsc{Rils}--\textsc{Rols}, able to find the right model on 57.84\% problems instances of the benchmark set \textsc{Feynman}; the same algorithm was even more impressive in solving problem instances from benchmark set \textsc{Strogatz}, begin successful in 83.57\% cases. The second best performing  approach is \textsc{AI-Feynman}, successful in solving 55.78\% and 27.14\% instances of the benchmark set \textsc{Feynman} and \textsc{Strogatz}, respectively. All other approaches are performing significantly worse, and none of them is able to solve more than 30\% instances of any of the two considered benchmark sets. 
	\item   Concerning the exact results when noise is presented with a value of 0.001, our \textsc{Rils}--\textsc{Rols} is still performing reasonably well, having found an exact model in more than 42\%, considering problems from the both benchmark set. Interestingly, \textsc{AI-Feynman} is quite sensible w.r.t.\ the change in noise, as its performances rapidly drop off -- it is successful on just 2.81\% of the all problem instances. Better from \textsc{AI-Feynman} are in this case \textsc{Dsr}, \textsc{Gplearn}, and \textsc{Afp-Fe} in this order, whose percentages of success range from about 15--17\%. 
	\item  Concerning the exact results when noise is high, that is with a value of 0.01,  our \textsc{Rils}--\textsc{Rols} is still reasonably good, having found an exact model for  34.77\% of the all problem instances. The second best is \textsc{Gplearn} which percentage of success is 12.75\%, followed by \textsc{Dsr} which is successful on just 9.23\% of the all problem instances. 
	\item \fxnote{Marko: opisati rezultate za $R^2 >0.999$}
\end{itemize}

>>>>

%Note that when the algorithm terminates with $R^2=1$ and $RMSE=0$, the obtained model and the exact model are always match (but also checked manually). However, in case when the algorithms deliver an error $R^2 \geq 0.999$, but $R^2 \neq 1$, these cases are all  manually checked by comparing weather the produced model match the exact model.

The exact results, thus the number of instances (in percentages) for which the output of respective algorithm match to the exact model, are displayed in the second block of each of the tables.

The numbers of instances (in percentages) for which each algorithms deliver a precise enough model (i.e. satisfying $R^2\geq 0.999$ upon its termination) are displayed in the third block of each of the tables.

@@

\begin{itemize}
	\item Concerning the exact results in case noise is not presented, the best performing algorithm is our \textsc{Rils}--\textsc{Rols}, able to find the right model on 57.84\% problems instances of the benchmark set \textsc{Feynman}; the same algorithm was even more impressive in solving problem instances from benchmark set \textsc{Strogatz}, begin successful in 83.57\% cases. The second best performing  approach is \textsc{AI-Feynman}, successful in solving 55.78\% and 27.14\% instances of the benchmark set \textsc{Feynman} and \textsc{Strogatz}, respectively. All other approaches are performing significantly worse, and none of them is able to solve more than 30\% instances of any of the two considered benchmark sets. 
	\item   Concerning the exact results when noise is presented with a value of 0.001, our \textsc{Rils}--\textsc{Rols} is still performing reasonably well, having found an exact model in more than 42\%, considering problems from the both benchmark set. Interestingly, \textsc{AI-Feynman} is quite sensible w.r.t.\ the change in noise, as its performances rapidly drop off -- it is successful on just 2.81\% of the all problem instances. Better from \textsc{AI-Feynman} are in this case \textsc{Dsr}, \textsc{Gplearn}, and \textsc{Afp-Fe} in this order, whose percentages of success range from about 15--17\%. 
	\item  Concerning the exact results when noise is high, that is with a value of 0.01,  our \textsc{Rils}--\textsc{Rols} is still reasonably good, having found an exact model for  34.77\% of the all problem instances. The second best is \textsc{Gplearn} which percentage of success is 12.75\%, followed by \textsc{Dsr} which is successful on just 9.23\% of the all problem instances.

	\item When comparing performances of the algorithms in terms of number of instances for which the algorithm's termination led to $R^2 \geq 0.999$ where no noise is integrated, our \textsc{Rils}-\textsc{Rols} is successful with 83.38\%. However, in contrast with the exact model percentages form above, there are a few methods delivering better percentages of success than ours, \textsc{Mrgp}, and \textsc{Operon}, with 92.69\%, and 86.93\%. In the other hand, as the  exact model percentage rates of these two approaches is rather low (0\% and 16\%) this indicates on a high level of over-fitting  of these algorithms. Thus, they are able to obtain a model that is accurate enough on the training data, but the produced model is wrong.

   \item  If comparing the performances of the competitors in terms of number of instances for which the algorithm's termination led to $R^2 \geq 0.999$ where the level of noise is 0.001, our \textsc{Rils}--\textsc{Rols} method has 78.62\% of a success rate. \textsc{Operon}, only has a better percentage rate of success among the other competitors, which is 83.08\%. However, the over-fitting is conspicuous as this algorithm is rarely able to deliver the model which fits ot the exact one (just in 0.31\%).

   \item  If comparing the performances of the competitors in terms of number of instances for which the algorithm's termination led to $R^2 \geq 0.999$ where the level of noise is 0.01, our \textsc{Rils}--\textsc{Rols} method is still highly accurate, delivering  62.92\% of a success rate. Interestingly, it is the best performing in terms of the percentage rate among other competitors. Just for the information, the second best is \textsc{Dsr}, with 16.92\% of a success rate. It can be concluded that our \textsc{Rils}--\textsc{Rols} method does not pay a big price in presence of a higher noise, like the other methods do. It still delivers high-accuracy models for the noisy data, but also many which match to the exact solutions (models).

\end{itemize}
